---
title: "Multiple-case preprocessing & analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple-case analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
pre {
  margin-top: 0.8rem;
  margin-bottom: 1.6rem;
}

.sourceCode {
  margin-bottom: 1.6rem;
}
```
## R code

### 1. Load the HARMONIC R package
```{r, results='hide'}
# install.packages("remotes")
# remotes::install_github("ERASMUSlab/HARMONIC")

# For exact reproducibility, pin clusterProfiler to the validated version:
# install.packages("./clusterProfiler_4.14.6.tar.gz",
#                  repos = NULL,
#                  type  = "source")

library(haRmonic)
packageVersion("clusterProfiler")
```

```{r, echo=FALSE, comment=""}
# install.packages("remotes")
# remotes::install_github("ERASMUSlab/HARMONIC")

# For exact reproducibility, pin clusterProfiler to the validated version:
# install.packages("./clusterProfiler_4.14.6.tar.gz",
#                  repos = NULL,
#                  type  = "source")

library(haRmonic)
packageVersion("clusterProfiler")
```
### 2. Preprocessing & Analysis

`HARMONIC_ANALYSIS()` runs the full preprocessing and enrichment workflow in a
single call. The main parameters are:

- `filepath` — path to the HARMONIC project directory (the `pwd` result from setup)
- `type` — analysis mode: `"standard"` (p < 0.05, fold change ≥ 2) or `"broad"` (p < 0.05, fold change ≥ 1.5)
- `full_condition` — all experimental conditions, in order
- `number_of_rep` — number of replicates per condition (must match `full_condition` in length)
- `DEG_list_name` — file name of the input DEG list
- `mango_design` — the comparison(s) to analyze, named as `A_B_UP` for genes up in A vs B (e.g. `"DAY4_DAY0_UP"`)
- `core` — number of CPU cores to use (increase for faster runs on multi-core machines)
- `ref_genome` — reference genome (`"mm"` for mouse, `"hs"` for human)
- `PASSED_RATIO` — minimum hit ratio (%) for an active tree to pass
- `PASSED_NUM` — minimum number of hit terms for an active tree to pass
- `similarity` — similarity threshold (%) for merging redundant terms
- `FC` — fold-difference threshold in HWES scores between cases; terms whose HWES differs by at least this factor across conditions are flagged as condition-specific (vs. commonly shared)
- `preprocessing` — whether to run preprocessing (`"T"` to run, `"F"` to skip)

```{r, eval=FALSE}
HARMONIC_ANALYSIS(filepath = "../HARMONIC_DIFF/HARMONIC/",
                  type = "standard",
                  full_condition = c("DAY0","DAY4","DAY7","DAY10","DAY14","DAY21"),
                  number_of_rep = c(3, 3, 3, 6, 3, 3),
                  DEG_list_name = "input_DEG_list.txt",
                  mango_design = c("DAY4_DAY0_UP","DAY7_DAY4_UP","DAY10_DAY7_UP","DAY14_DAY10_UP","DAY21_DAY14_UP"),
                  core = 5,
                  ref_genome = "mm",
                  PASSED_RATIO = 20,
                  PASSED_NUM = 2,
                  similarity = 70,
                  FC = 2,
                  preprocessing = "T")
```

### 3. Modify parameters

Key parameters can be adjusted depending on your design. For example, to apply
stricter filtering and relaxed redundancy merging:

- `PASSED_RATIO = 25` — require a higher hit ratio per active tree
- `PASSED_NUM = 3` — require more hit terms per active tree
- `similarity = 50` — merge terms at a lower similarity threshold
- `FC = 1.5` — relax the HWES fold-difference threshold for calling condition-specific terms
- `preprocessing = "F"` — skip preprocessing if already run

```{r, eval=FALSE}
HARMONIC_ANALYSIS(filepath = "../HARMONIC_DIFF/HARMONIC/",
                  type = "standard",
                  full_condition = c("DAY0","DAY4","DAY7","DAY10","DAY14","DAY21"),
                  number_of_rep = c(3, 3, 3, 6, 3, 3),
                  DEG_list_name = "input_DEG_list.txt",
                  mango_design = c("DAY4_DAY0_UP","DAY7_DAY4_UP","DAY10_DAY7_UP","DAY14_DAY10_UP","DAY21_DAY14_UP"),
                  core = 5,
                  ref_genome = "mm",
                  PASSED_RATIO = 25,   # changed from 20
                  PASSED_NUM = 3,      # changed from 2
                  similarity = 50,     # changed from 70
		  FC = 1.5,            # changed from 2
                  preprocessing = "F") # changed from "T"
```


### 4. Dynamic analysis

Dynamic analysis identifies active trees that are **commonly detected across a
specified range of conditions**, helping reveal biological processes that are
consistently active over a trajectory (e.g. a differentiation time course)
rather than appearing in only one comparison.

- `condition = c(1, 3)` — the range of conditions (by index) to scan for commonly detected trees
- `dynamic_analyㅑsis = "T"` — enable dynamic analysis (`"F"` to disable)

```{r, eval=FALSE}
HARMONIC_ANALYSIS(filepath = "../HARMONIC_DIFF/HARMONIC/",
                  type = "standard",
                  full_condition = c("DAY0","DAY4","DAY7","DAY10","DAY14","DAY21"),
                  number_of_rep = c(3, 3, 3, 6, 3, 3),
                  DEG_list_name = "input_DEG_list.txt",
                  mango_design = c("DAY4_DAY0_UP"),
                  core = 5,
                  ref_genome = "mm",
                  PASSED_RATIO = 25,   
                  PASSED_NUM = 3,      
                  similarity = 50,     
                  FC = 1.5,           
                  condition = c(1,3),
                  dynamic_analyisis = "T", 
                  preprocessing = "F")
```

